Data Munging

### 2019-12-13 scripts/dataMunge.Rmd
(1) Fixed error in AREA_HA calculation. (2) Used more efficient dplyr cut function for size_cat factor.

The survey design is stratified by State.  Within each state, an unequal probability category based on 5 lake size classes is used.  As of 12/12/2019, NLA has not provided data on waterbody size.  Size was estimated by doing a spatial overlay of the waterbody lat/lon with NHDPlusV2.  Surface area (AREASQKM) was then extracted from NHDPlusV2.  The data currently contains `r dg %>% as.data.frame() %>% filter(is.na(AREASQKM)) %>% {nrow(.)}` missing waterbody size values.  For now, we'll use these data to add the size categories to the data file.
```{r}
dg <- dg %>%
  mutate(
    # add state
    state = substr(site.id, 7, 8),
    AREA_HA = AREASQKM / 100, # 100Ha in 1 km2 
    size_cat = ifelse(AREA_HA <= 4,
                      "LE4",
                      ifelse(AREA_HA > 4 & AREA_HA <= 10,
                             "G4_LE10",
                             ifelse(AREA_HA > 10 & AREA_HA <= 20,
                                    "G10_LE20",
                                    ifelse(AREA_HA > 20 & AREA_HA <= 50,
                                           "G20_LE50",
                                           ifelse(AREA_HA > 50,
                                                  "G50",
                                                  NA))))))

@@

Concentrations of CO~2~, CH~4~, and N~2~O in ambient air over the waterbodies were measured and used to calculate the equilibrium dissolved gas concentration (variables sat.co2, sat.ch4, and sat.n2o).  These data are not needed in subsequent analysis and will be removed from the dataframe.
```{r}
dg <- dg %>% filter(sample.source == "DG") # select dissolved gas 'DG' samples
```

>>>>

The survey design is stratified by State.  Within each state, an unequal probability category based on 5 lake size classes is used.  As of 12/12/2019, NLA has not provided data on waterbody size.  Size was estimated by doing a spatial overlay of the waterbody lat/lon with NHDPlusV2.  Surface area (AREASQKM) was then extracted from NHDPlusV2.  The data currently contains `r dg %>% as.data.frame() %>% filter(is.na(AREASQKM)) %>% {nrow(.)}` missing waterbody size values.  For now, we'll use these data to add the size categories to the data file.
```{r}
dg <- dg %>%
  mutate(
    # add state
    state = substr(site.id, 7, 8),
    AREA_HA = AREASQKM * 100, # 100Ha in 1 km2 
    size_cat = factor( cut( AREASQKM,
                            breaks = c( -Inf, 0.04, 0.1, 0.2, 0.5, Inf ),
                            labels = c( "min_4", "4_10", "10_20", "20_50", "50_max") ) )
    )

@@
